---
title: "Homework Tidyverse"
subtitle: "Zeynep Yilmaz"
format:
  html:
    embed-resources: true
editor: visual
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  eval = TRUE,
  fig.height = 3,
  fig.width = 6,
  fig.align = "center"
)

ggplot2::theme_set(ggplot2::theme_bw())
```

```{r}
#| message: false
#| warning: false
if(!require("pacman")) install.packages("pacman")
pacman::p_load(tidyverse, gapminder, broom)
```

# Problem 1:

Copy paste and run the tribble given below.

```{r}
tribble( ~x,    ~y,    ~w,    ~z,
              210,   300,   220,   180,
              102,   100,   119,   187,
              176,   175,   188,   173,
              87,    95,   91,     94,
              202,   210,  234,    218,
              110,   122,  131,    128,
) -> dt
```

## 1_a:

- Use and show a `map` function to find the "mean" of each column of the `dt` data table
Using mean inside map.

```{r}
map(dt,mean)
```

## 1_b:

- Use and show a `map` function to find the "standard deviation" of each column of the `dt` data table.
Using sd in map.

```{r}
map(dt,sd)
```

## 1_c:

- Use and show a map function that will calculate the "square root" of each value of each column of the data table `dt`.
Using sqrt inside map().

```{r}
map(dt,sqrt)
```

## 1_d:

- Use R code to find the "mean", "max", "1st Quartile", "3rd Quartile", "Median", and "Min" for each column of the `dt` data table. (Hint: You do not have to use a map function)

```{r}
summary(dt)
```

# Problem 2:

Write a function that uses a for loop that, for each iteration, randomly draws 5 observations from an exponential distribution with "rate" parameter 1 (use `rexp()`) and calculates its "mean". It should do this 10,000 times. Choose an appropriate plot to plot the distribution of "means".

- Solution: We want to fill a vector with values. We should create this vector beforehand using the `vector()` function.

- We want to create a vector of size 10 and we want its values to be numeric.

- Then we want to randomly draw 5 observations from an exponential distribution with "rate" parameter 1 (use `rexp(5, rate = 1)`) and calculates its "mean"

- Now we want to create a vector that has length of 10 and its values are mean of this randomly selected exponential functions

- Now we want to draw the histogram. But we need to have a data frame. Let us define a data frame.

- Now we can draw a histogram

  - **Now, let's solve the problem all in one step**

## 2_a:

- Repeat part 1 by using a `map_*()` function.
Using map_dbl since we want one decimal numeric mean for each iteration.
1:10000 is the for loop using map.
Using rexp() to choose 5 random observations with rate = 1.

```{r}
means <- map_dbl(1:10000, ~ mean(rexp(5, rate = 1)))
means
```

## 2_b:

- Repeat part 1 by using the `replicate()` function.
Replicate iretates 10000 times so we didn't need 1: and ~ since it gives back a list but we don't need a list for replicate.

```{r}
means_2 <- replicate(10000, mean(rexp(5, rate = 1)))
means_2
```
## 2_c:

- Use a another for loop that will print out plots for sample sizes of 5, 10, and 20 observations (instead of just 5).

Using replicate with a general for loop putting the random selection numbers of 5,10,20 with mean and rexp. Then creating a histogram with hist(calculation, titles).

```{r}
for (n in c(5, 10, 20)) {
  
  means <- replicate(10000, mean(rexp(n, rate = 1)))
  
  hist(means,
       main = paste("Sample Size =", n),
       xlab = "Sample Means")
}
```

# Problem 3:

- Use and show R coding to calculate the "standard deviation" for each variable of the data table `mtcars` using the “Special For Loop Method”.

**Note:**

- There are two methods. Please see a quick explanation for both methods in below:

  - Method 1:

```{r}
  multi_return <- function() {
    my_list <- list("color" = "red", "size" = 20, "shape" = "round")
    return(my_list) 
  }
multi_return()
```

```{r}
a <- multi_return()
a$color

a<- multi_return()
a$shape

a<- multi_return()
a$size
```

- Method 2. The "Special For Loop Method"

```{r}
#| eval: false
output <- vector("double", ncol(mtcars))  # 1. output (use appropriate data frame instead of df)
for (i in seq_along(mtcars)) {            # 2. sequence
  output[[i]] <- median(mtcars[[i]])      # 3. body
}
output
```

- **Solution:** coding to calculate the "standard deviation" for each variable

# Problem 4:

Generate a sequence of dates from **January 1, 2015** to **December 31, 2025**, spaced by every **two months**. Extract the **year**, **quarter**, and **ISO week number** for each date.

```{r}
library(lubridate) #for date package

dates <- seq( #sequencing the wanted
  from = as.Date("2015-1-1"),
  to = as.Date("2025-12-31"),
  by = "2 months"
)

year <- year(dates) #identifying and extracting the values
quarter <- quarter(dates)
iso_week <- isoweek(dates)

date_data <- data.frame( #making it into a dataframe
  date = dates,
  year = year(dates),
  quarter = quarter(dates),
  iso_week = isoweek(dates)
)

date_data
```

# Problem 5

Given the following dates, compute the difference in **months** and **weeks** between each consecutive pair.

``` {r}
library(lubridate) #for as.Date, interval, from to by.

sample_dates <- c("2018-03-15", "2020-07-20", "2023-01-10", "2025-09-05")

month_diff <- time_length(
  interval(sample_dates[-length(sample_dates)], sample_dates[-1]), #reverse
  "months"
)

week_diff <- time_length(
  interval(sample_dates[-length(sample_dates)], sample_dates[-1]), 
  "weeks"
)
# Put results into a data frame
date_differences <- data.frame( #from reverse list to nromal list
  from = sample_dates[-length(sample_dates)],
  to = sample_dates[-1],
  months = month_diff,
  weeks = week_diff
)

date_differences
```

# Problem 6

Given a list of mixed date formats, use `map()` and `possibly()` from `purrr` to safely convert them to **Date** format and extract the **month name**.

``` {r}
library(tidyverse)
library (lubridate)

date_strings <- list("2023-06-10", "2022/12/25", "15-Aug-2021", "InvalidDate")

safe_date <- possibly( #possibly to not make the code crash by invaliddate
  function(x) {
    parse_date_time(x, orders = c("ymd", "dmy"))
  },
  otherwise = NA
)

dates <- map(date_strings, safe_date) #converts date string into writing

month_names <- map(dates, ~ month(.x, label = TRUE)) #extracts the month from each date 
month_names
```
